\documentclass[11pt,a4paper]{report} 

% Für doppelseitigen Ausdruck (nur bei > 60 Seiten sinnvoll)
% \usepackage{ifthen}
% \setboolean{@twoside}{true}
% \setboolean{@openright}{true} 

\include{preamble} % alle Pakete und Einstellungen

% Hier anpassen 
%\newcommand{\welchethesis}{Bachelor}
\newcommand{\welchethesis}{Master}
\newcommand{\thesisofwas}{of Science}
%\newcommand{\studiengang}{Technische Informatik}
% \newcommand{\studiengang}{Medizintechnik}
\newcommand{\studiengang}{Informationstechnik}
\newcommand{\titel}{Compositional Gaussian Splatting for Multi-Object and Dynamic Scene
Synthesis}
\newcommand{\kurztitel}{Compositional Gaussian Splatting}
\newcommand{\autor}{Luca Hebeda}
\newcommand{\datum}{31. Oktober 2025} % Abgabedatum
\newcommand{\ort}{Mannheim}
\newcommand{\hochschullehrer}{Prof.\ Dr.\ Marcus Vetter}
\newcommand{\zweikorrektor}{M. Sc. Yannick Bukschat}

\begin{document}
\include{vorspann} % Titelseite, Erklärungen, etc.

\begin{abstract}
We present a modular and methodologically transparent pipeline for multi-view and temporally consistent synthetic dataset generation based on Gaussian Splatting. 
The framework integrates static 3D Gaussian Splatting (3DGS) and persistent Dynamic 3D Gaussian Splatting (D3DGS) into a unified system that enables reconstruction, export, and recomposition of dynamic multi-object scenes. 
Using a hardware-synchronized 56-camera rig, we capture spatially dense and temporally aligned sequences that serve as the basis for static background reconstruction and dynamic object modeling. 
A dedicated compositing stage allows the exported Gaussian models to be reinserted, duplicated, or combined into novel virtual environments, producing multi-view renderings with synchronized RGB, depth, segmentation, and occlusion annotations. 
Additional modules provide pose estimation, keypoint propagation, and temporal supersampling for motion-consistent supervision signals. 
Qualitative evaluations demonstrate that the pipeline maintains spatial coherence, robust occlusion reasoning, and temporal stability across views and environments. 
The system establishes a practical foundation for controllable 4D dataset generation and for benchmarking methods in dynamic scene understanding.
\end{abstract}


\tableofcontents

\onehalfspacing

% \linenumbers
\input{Kapitel/1_Introduction} % Externe Datei einbinden
\input{Kapitel/2_Fundamentals.tex}
\input{Kapitel/2_State_of_the_Art} % Externe Datei einbinden
\input{Kapitel/3_Methodology} % Externe Datei einbinden
\input{Kapitel/4_Results} % Externe Datei einbinden
%\input{Kapitel/5_Comparison} % Externe Datei einbinden
\input{Kapitel/5_Conclusion} % Externe Datei einbinden
%\input{kapitel/kapitel5} % Externe Datei einbinden


\newpage

% Listen wenn überhaupt ans Ende und nicht an den Anfang.
% Meist ist das aber unnötig.
% \listoffigures % Liste der Abbildungen 
% \listoftables % Liste der Tabellen
% \newpage

\bibliographystyle{plain} % Literaturverzeichnis
% \bibliography{references}

\begin{btSect}{thesis} % mit bibtopic Quellen trennen
\addcontentsline{toc}{chapter}{Literaturverzeichnis}
\section*{Literaturverzeichnis}
\btPrintCited
\end{btSect}

% dann mit "bibtex thesis1" und "bibtex thesis2" arbeiten

\end{document}
;;; Local Variables:
;;; ispell-local-dictionary: "de_DE-neu"
;;; End:
